---
title: "Geographic_ABM_on_Impacts_of_NPIs_on_the_Vaccination_Process_In_Kenya"
author: "Vincent Okoth"
date: "11/23/2021"

knit: (function(input_file, encoding) {
  out_dir <- 'D:\GEGIS\Project\R\agent_based_modelling_of_COVID_19_NPIs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background

The devastating impact of the COVID-19 pandemic on global health and economics 
has ushered in a period of fast vaccine research, development, manufacturing, and distribution. With several vaccines currently available for use in Kenya, this
markdown file outlines the details of the Geographic agent-based model on 
impacts of NPIs (None pharmaceutical interventions) and policy changes on the 
Vaccination Process In Kenya. The model implements agents that undergo vaccination
as they progress through various stages of the COVID-19 disease. The agents 
have various psychological states to allow the exploration of interactions, 
compliance, and much more. A geographical network is defined, with a social network
that governs agent's interaction patterns over the network. 

## Developing the agent

An agent can have two states: a psychological state, and a biological state.
Some examples of the biological state include 'Quarantine', social distancing, 
and believing in a government vaccination directive. Some of these states will
be explored later in the model.

The biological states have several specific cases including:

1. Susceptible
2. Exposed
3. Infectious
4. Recovered
5. Susceptible and vaccinated
6. Exposed and vaccinated
7. Infectious and vaccinated
8. Post COVID immune
9. Naturally Immune
10. Death

Whereas existing SIR models break #3 into 2 to three stages including hospitalization
and death,including the stages will require that we define more parameters that
will complicate the model. Initially, the agents could be defined just by their
biological states, with the assumption that most people are in state 1. Others
would be in state 8 or 9 already which are more of the same. 

Global variables will be used to define the model, and the states defines as 
named in the above list.



```{r, cache=TRUE}
library(gganimate)
library(ggplot2)
library(dplyr)
library(sna)
library(knitr)
library(reshape2)
library(igraph)

States <<- 10



Statenames <<- c("Susceptible",
                 "Exposed",
                 "Infectious",
                 "Recovered",
                 "Susceptible and vaccinated",
                 "Exposed and vaccinated",
                 "Infectious and vaccinated",
                 "Post COVID immune",
                 "Naturally Immune",
                 "Death")



Statelabels <<- c("Susceptible",
                 "Exposed",
                 "Infectious",
                 "Recovered",
                 "Susceptible \n& vaccinated",
                 "Exposed \n& vaccinated",
                 "Infectious \n& vaccinated",
                 "Post COVID immune",
                 "Naturally Immune",
                 "Death")



age_group <- c("13 to 75+")



makeAgent <- function(psychstate, biostate, age_group) 
{
  return(list(psychstate = psychstate,
              biostate = biostate,
              age_group = age_group))
}
print(makeAgent(1,2,3))
```
## Biological model timecourse

Once an agent is infected or vaccinated, it is assumed that the trajectory of the
disease or vaccination state is fixed and progresses, eventually leading to recovery,
post COVID-19 immunity, or death. The model therefore needs a way to automatically
transition the biological state in a reasonable way, which is more of a task model
that keeps tract of time various states. This considers only the natural progression
of the disease and vaccination process. Thus, a time course with a 1-day unit is
modeled. 



```{r, cache=TRUE}
makeAgent <- function(psychstate, biostate, age_group) 
{
  return(list(psychstate = psychstate,
              biostate = biostate,
              age_group = age_group,
              nextbiostate=NA,
              biostatecountdown=NA))
}
print(makeAgent(1,2,3))
```



Despite the assumption that most agents are in biostate 1, a function is required
to update the agents' states everyday, change the countdown, and transition agents
to next states if needed. If there's no next state (Post COVID immune, Naturally Immune,
or Death), the state is set to NA and no longer updated. 



```{r, cache=TRUE}
updateAgent <- function(agent)
{
  agent$biostatecountdown <- agent$biostatecountdown -1
  if(agent$biostatecountdown <=0)
  {
    agent$biostate <- agent$nextbiostate
    agent$biostatecountdown <- NA
  }
  return(agent)
}
```



## Creating transition matrix
With a way of transitioning the agent now in place, a set of possible transitions
need to be implemented and the timing of each state defined. To simplify the model, 
all timing distributions are made with a minimum and maximum parameter for each
state, making more of a markov process. Two pathways can be programmed through 
the stages, with several branch points such as (possibility of recovering after acquiring COVID-19, death vs recovery, vaccination vs post COVID-19 immunity, and so on). The
progression of the disease and vaccination process is completely defined by the
data in the transition matrix, with a generic function that automatically updates
the state of each agent every day. 



```{r, cache=TRUE}
# * 1. Susceptible
# * 2. Exposed
# * 3. Infectious
# * 4. Recovered
# * 5. Susceptible and vaccinated
# * 6. Exposed and vaccinated
# * 7. Infectious and vaccinated
# * 8. Post COVID immune
# * 9. Naturally Immune
# * 10. Death


bioTransition <- matrix(0,States,States)
bioMin <- matrix(1,States)      #state time minimum
bioMax <- matrix(1,States)      #state time maximum



bioMin[2] <- 2            #Exposed and unvaccinated/vaccinated for 2 to 14 days
bioMax[2] <- 14
bioTransition[2,3] <- .5  #transition to infectious
bioTransition[2,6] <- .5  #transition to infectious and vaccinated

bioMin[3] <- 14           #It takes 2 to 6 weeks for an infectious agent to recover 
bioMax[3] <- 42 
bioTransition[3,4] <- .94
bioTransition [3,10] <- .06

bioMin[6] <- 14           #Transition from exposed and vaccinated to recovery/death
bioMax[6] <- 42
bioTransition[6,4] <-.97
bioTransition[6,10] <- .03

bioMin[5] <- 14           #Transition from 5 to 6/7/8
bioMax[5] <- 42          
bioTransition[5,6] <- .03 
bioTransition[5,7] <- .03
bioTransition[5,8] <- .94

bioMin[7] <-14          #Transition from 7 to 4/8/10
bioMax[7] <-42
bioTransition[7,4] <- .48
bioTransition[7,8] <- .48
bioTransition[7,10] <- 04

bioMin[4] <-125
bioMax[4] <- 365
bioTransition[4,8] <- .5 #Transition from 4 to 8/1
bioTransition[4,1] <- .5



setAgentState<- function(agent, biostate)
{
  agent$biostate <- biostate
  if(sum(bioTransition[biostate,])>0) # this state transitions to something else.
  {
    ##which state do we go to?
     agent$biostatecountdown <- sample(x=seq(bioMin[biostate],bioMax[biostate]),1) #how long will we state in this state?
     agent$nextbiostate <- sample(1:STATES, prob=bioTransition[agent$biostate,],size=1)
     
  } else{
   agent$biostatecountdown <- NA
   agent$nextbiostate <- NA   ##just so we can tell if the agent is finished.
  }
  return(agent) 
}


transitionAgent<- function(agent)
{
   return(setAgentState(agent,agent$nextbiostate))
}

updateAgent<- function(agent)
{
  if(!is.na(agent$biostatecountdown))
  {
   agent$biostatecountdown <- agent$biostatecountdown -1
    if(agent$biostatecountdown <=0)  ##new state
    {
       agent <- transitionAgent(agent)

    }
  }
   return(agent)
}
```




```{r, cache=TRUE, fig.width=8, fig.height=8}
par(xpd=NA,mar=c(5,5,8,5))
gplot(bioTransition,label=Statelabels)
```


