---
title: "Geographic_ABM_on_Impacts_of_NPIs_on_the_Vaccination_Process_In_Kenya"
author: "Vincent Okoth"
date: "11/23/2021"

knit: (function(input_file, encoding) {
  out_dir <- 'D:\GEGIS\Project\R\agent_based_modelling_of_COVID_19_NPIs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background

The devastating impact of the COVID-19 pandemic on global health and economics 
has ushered in a period of fast vaccine research, development, manufacturing, and distribution. With several vaccines currently available for use in Kenya, this
markdown file outlines the details of the Geographic agent-based model on 
impacts of NPIs (None pharmaceutical interventions) and policy changes on the 
Vaccination Process In Kenya. The model implements agents that undergo vaccination
as they progress through various stages of the COVID-19 disease. The agents 
have various psychological states to allow the exploration of interactions, 
compliance, and much more. A geographical network is defined, with a social network
that governs agent's interaction patterns over the network. 

## Developing the agent

An agent can have two states: a psychological state, and a biological state.
Some examples of the biological state include 'Quarantine', social distancing, 
and believing in a government vaccination directive. Some of these states will
be explored later in the model.

The biological states have several specific cases including:

1. Susceptible
2. Exposed
3. Infectious
4. Recovered
5. Susceptible and vaccinated
6. Exposed and vaccinated
7. Infectious and vaccinated
8. Post COVID immune
9. Naturally Immune
10. Death

Whereas existing SIR models break #3 into 2 to three stages including hospitalization
and death,including the stages will require that we define more parameters that
will complicate the model. Initially, the agents could be defined just by their
biological states, with the assumption that most people are in state 1. Others
would be in state 8 or 9 already which are more of the same. 

Global variables will be used to define the model, and the states defines as 
named in the above list.



```{r, cache=TRUE}
library(gganimate)
library(ggplot2)
library(dplyr)
library(sna)
library(knitr)
library(reshape2)
library(igraph)

States <<- 10



Statenames <<- c("Susceptible",
                 "Exposed",
                 "Infectious",
                 "Recovered",
                 "Susceptible and vaccinated",
                 "Exposed and vaccinated",
                 "Infectious and vaccinated",
                 "Post COVID immune",
                 "Naturally Immune",
                 "Death")



Statelabels <<- c("Susceptible",
                 "Exposed",
                 "Infectious",
                 "Recovered",
                 "Susceptible \n& vaccinated",
                 "Exposed \n& vaccinated",
                 "Infectious \n& vaccinated",
                 "Post COVID immune",
                 "Naturally Immune",
                 "Death")



age_group <<- c("13 to 75+")



makeAgent <- function(psychstate, biostate, age_group) 
{
  return(list(psychstate = psychstate,
              biostate = biostate,
              age_group = age_group))
}
print(makeAgent(1,2,3))
```
## Biological model timecourse








